---
title: "trying lasso and stuff again"
author: "Eleanor Noble"
date: "May 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
library(ggmap)
library(RColorBrewer)
library(viridis)
library(plotly)
library(maps)
library(lubridate)
library(plotly)
library(janitor)
library(dplyr)
library(sf)
library(caret)
```

Hopefully this code works for reading in all the files
```{r}
#ev.mpls.raw this is all the rental license properties
rental_prop_total<-read_sf("./rental_license_census_data.shp")%>%janitor::clean_names()
census_pop<-read.csv("./census2010pop.txt")
parcels_2018<-read.csv("./parcel_2018_post_cleaning.csv")
ev.mpls<-read.csv("./rental_props_parcel_eviction_join.csv")

census_pop$GEOG_UNIT<-as.numeric(census_pop$GEOG_UNIT)
rental_prop_total$geoid10<-as.numeric(rental_prop_total$geoid10)

rental_prop_total<-rental_prop_total%>%
  left_join(census_pop,by=c("geoid10"="GEOG_UNIT"))

parcels_2018<-parcels_2018%>%
  select(pin,emv_total,year_built,homestead)%>%
  mutate(apn = substr(pin,5,17) )

parcels_2018$apn<- as.numeric(parcels_2018$apn)
rental_prop_total$apn<-as.numeric(rental_prop_total$apn)

rental_prop_total<-rental_prop_total%>%
  left_join(parcels_2018,by=c("apn"="apn"))

ev.mpls<-ev.mpls%>%
  select(apn,eviction_count)

rental_prop_total<-rental_prop_total%>%
  left_join(ev.mpls, by=c("apn"="apn"))

rental_prop_total$eviction<-ifelse(is.na(rental_prop_total$eviction_count)==TRUE,0,1)

# write_csv(rental_prop_total,"log_reg_clean.csv")
rentals_evics<-read_csv("./log_reg_clean.csv")%>%janitor::clean_names()

rentals_evics<-rentals_evics%>%
  mutate(eviction = as.factor(eviction))

ev.mpls<-ev.mpls.raw%>%
  janitor::clean_names()%>%janitor::remove_empty_rows()%>%
  select(-(1:14))%>%
  select(-(16:26))

# write_csv(ev.mpls,"./to_clean.csv")

ev.mpls<-read.csv("./rental_props_parcel_eviction_join.csv")
rental_license_ct_data<-read_sf("./rental_license_census_data.shp")

census_pop<-read.csv("./census2010pop.txt")

ev.mpls<-ev.mpls%>%
  left_join(rental_license_ct_data,by=c("licenseNum"="licenseNum"))

census_pop$GEOG_UNIT<-as.numeric(census_pop$GEOG_UNIT)
ev.mpls$GEOID10<-as.numeric(ev.mpls$GEOID10)

ev.mpls<-ev.mpls%>%
  left_join(census_pop,by=c("GEOID10"="GEOG_UNIT"))

# write_csv(ev.mpls,"to_clean_part3.csv")

ev.mpls2<-read_csv("./to_clean_part4.csv")%>%janitor::clean_names()

ev.mpls2<-ev.mpls2%>%
  filter(!is.na(license_num))

rownames(ev.mpls2)<-ev.mpls2$license_num

ev.mpls2<-ev.mpls2%>%
  select(-license_num)
```

this is setting up the logistic regression 
```{r}
train_ctrl_cv10 <- trainControl(method = "cv", number = 10)

set.seed(333)
evic_mod <- train(
    eviction ~ .,
    data = rentals_evics,
    method = "glm",
    family = binomial,
    metric = "Accuracy",
    trControl = train_ctrl_cv10,
    na.action = na.omit
)

evic_mod
evic_mod$results
```
This should show the results of the confusion matrix for logistic regression
```{r}
rentals_evics[is.na(rentals_evics)] <- 0
# predict() below uses 0.5 as a probability threshold by default
mod_pred <- predict(evic_mod, newdata = rentals_evics)

# Compute confusion matrix and statistics
conf_mat <- confusionMatrix(data = mod_pred, reference = rentals_evics$eviction, positive = "1")
# Print results
conf_mat
```

This shows that the linear reg assumptions hold alright (no qq plot tho because that was kinda fuked up). This means using lasso for shrinkage is probably alright because there aren't any huge violations
```{r}
mod1_reg <- data.frame(
    residual = residuals(lasso_var_mod),
    predicted = fitted.values(lasso_var_mod)
)

# Plot
ggplot(mod1_reg, aes(x = predicted, y = residual)) +
    geom_point() +
    geom_smooth(color = "blue", lwd = 3) + # Add a smooth trend
    geom_hline(yintercept = 0, color = "red") # Add the y = 0 line

```

The OLS comparison model and the RMSE at the bottom to compare to other models.
```{r}
ev.mpls.scaled <- dist(scale(ev.mpls2, center = FALSE, scale = TRUE))

least_squares_mod <- lm(eviction_count ~ ., data = ev.mpls2)
coefficients(least_squares_mod)

summary(least_squares_mod)

train_ctrl_cv10 <- trainControl(method = "cv", number = 7)

set.seed(253)

mod1_cv10 <- train(
    eviction_count ~ .,
    data = ev.mpls2,
    method = "lm",
    trControl = train_ctrl_cv10,
    na.action = na.omit
)

mean(mod1_cv10$resample$RMSE^2)

```

All the lasso stuff
```{r}
train_ctrl_cv10 <- trainControl(method = "cv", number = 10)

# Create a grid of lambda values
lambdas <- 10^seq(-3, 3, length.out = 100)

# Fit LASSO models for all of the lambdas
set.seed(74)
lasso_mod <- train(
    eviction_count ~ .,
    data = ev.mpls2,
    method = "glmnet",
    trControl = trainControl(method = "cv", number = 7),
    tuneGrid = data.frame(alpha = 1, lambda = lambdas),
    metric = "RMSE",
    na.action = na.omit
)

# Model coefficients for lambda = 100
coefficients(lasso_mod$finalModel, 100)

# Plot summary of coefficient estimates
plot(lasso_mod$finalModel, xvar = "lambda", label = TRUE, col = rainbow(20))

# Zoom in
plot(lasso_mod$finalModel, xvar = "lambda", label = TRUE, col = rainbow(20), ylim = c(-.0001,.0001))

# What variables do the numbers correspond to?
rownames(lasso_mod$finalModel$beta)

plot(lasso_mod,xlim = c(-1,5))
plot(lasso_mod,xlim = c(-.1,1))

mean(lasso_mod$resample$RMSE^2)
```

picking the best lambdas
```{r}
best_lambdas <- function(model) {
    # Extract the results table
    res <- model$results
    # Extract the K in K-fold CV
    k <- model$control$number
    # Compute the standard error (SE) of the RMSE estimate
    res$rmse_se <- res$RMSESD/sqrt(k)
    # Which lambda resulted in the lowest RMSE?
    index_lambda_min <- which.min(res$RMSE)
    lambda_min <- res$lambda[index_lambda_min]
    # Compute 1 SE below and above the minimum RMSE
    res$rmse_lower <- res$RMSE - res$rmse_se
    res$rmse_upper <- res$RMSE + res$rmse_se
    rmse_lower <- res$RMSE[index_lambda_min] - res$rmse_se[index_lambda_min]
    rmse_upper <- res$RMSE[index_lambda_min] + res$rmse_se[index_lambda_min]
    res$within_1se <- res$RMSE >= rmse_lower & res$RMSE <= rmse_upper
    index_lambda_1se <- max(which(res$within_1se))
    lambda_1se <- res$lambda[index_lambda_1se]
    p <- ggplot(res, aes(x = lambda, y = RMSE)) +
        geom_pointrange(aes(ymin = rmse_lower, ymax = rmse_upper))
    print(p)
    output <- res[c(index_lambda_min, index_lambda_1se),c("lambda", "RMSE")]
    rownames(output) <- c("lambda_min", "lambda_1se")
    output
}

lambda_choices <- best_lambdas(lasso_mod)
lambda_choices

# Coefficients for the lambda_1se LASSO model
lambda_choices["lambda_1se", "lambda"]

coefficients(lasso_mod$finalModel, .03)

#lasso mod rmse
mean(lasso_mod$resample$RMSE^2)

```

