---
title: "MN court data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
library(ggmap)
library(RColorBrewer)
library(viridis)
library(plotly)
library(maps)
library(lubridate)
library(plotly)
```


```{r}
courts_raw<-read_csv("./copy of extract 2.6.19.csv")%>%janitor::clean_names()
judgments_simply<-read_csv("./basic judgement descriptions.csv")%>%janitor::clean_names()
```

```{r functions}
# Custom functions

# Define LLCs and corporations with regular expressions

llc <- c("LLC\\.?|Inc\\.?|\\bL\\.?l{0,2}P\\.?\\b|Corp\\b|Company|Co\\.|\\bCo\\b|Ltd\\.?|\\bG\\.?P\\.?\\b|
         \\(GP\\)|\\sL\\sP|Partnership|Limited^|Corporation|Public Housing|
         Inc\\.?|Office|Apartments?|Bank\\b|Mortgage|State of Minnesota|
         Minnesota Housing|Properties|Trust\\b|Investments|Credit Union") 
    # Types of corporations or LLCs
    # Including Banks, Public Housing, Corps, Offices, etc.

# Function to capitalize each word, including common acronyms

simpleCap <- function(x) {
  
  llc.cap <- c("Llc\\.?|L\\.?l{0,2}p\\.?|G\\.?p\\.?|\\(gp\\)|Iii?|Us\\s|Fsb|Na$")
  
  s <- strsplit(tolower(x), "\\s")[[1]] # split if there is a space
  
  gsub("(-\\w)", "\\U\\1", # next in gsub to capitalze after hyphens post-hoc
       paste(toupper(substring(s, 1,1)), substring(s, 2), sep="", collapse=" "), 
       pe = TRUE) %>%
    str_replace_all(llc.cap, toupper) # capitalize common acronyms
}

# Custom function for turning name into separate columns (first, middle, last, suffix)

# Process: Input single variable, output multiple ones (based on variable name/abbreviation?)
# Screen for LLCs, use variable to differentiate, then use humaniformat to reverse, possibly clear up periods,
# then parse into columns with first column name (or first letters) as prefix

# Works whether data is Smith, John or John Smith and tries to differentiate companies and LLCs from people

# Usage: input list, turns into data frame with multiple columns (original, column plus separate name elements)

sep.names <- function(x, y) { # x for list with full unformatted names; y for name of list
  
  x.name <- y # extract name for later use, still working on this

  llc.cap <- c("Llc\\.?|L\\.?l{0,2}p\\.?|G\\.?p\\.?|\\(gp\\)|Iii?|Fsb|Us\\s") # acronyms to capitalize
  
  suffix <- c("Jr\\.?|Sr\\.?|1$|Iii?") # possible suffices
  
  x2 <- as.data.frame(x) %>% 
    filter(!is.na(x) | x == "") %>%
    mutate(x2 = sapply(x, simpleCap)) %>% # capitalize each word
    mutate(llc = ifelse(grepl(llc, x2), 1, 0)) %>%
    mutate(x2 = str_replace_all(x2, llc.cap, toupper)) %>% # capitalize LLCs, etc
    mutate(x2 = str_replace_all(x2, "\\(([^\\)]+)\\)", "")) %>% # remove parentheses
      # mutate(x = str_replace_all(x, "\\,\\s\\.", "")) %>% # remove trailing punctuation
    mutate(x2 = ifelse(llc == 0, format_reverse(x2), x2)) %>% # reverse when there are commas
    mutate(suffix = str_detect(x2, "^(.(?!(Jr\\.?|Sr\\.?|1$|Iii?)))+$")) %>%
    mutate(x2 = ifelse((llc == 0 & str_detect(x2, "^(.(?!(Jr\\.?|Sr\\.?|1$|Iii?\\b)))+$") == "FALSE"), 
                          # only if not llc and doesn't already have common suffix at the end
                       str_replace_all(x2, "(.*)\\s(Jr\\.?|Sr\\.?|1$|Iii?)\\s(.*)",
                                                "\\ 1 \\3 \\2"), x2)) # Move suffix to the end (after reversing)
  
  parse <- x2 %>%
    filter(llc == 0) %>%
    do(parse_names(.$x2)) %>%
    select(-salutation) %>%
    rename_at(vars(-full_name), function(x) paste0(x.name, "_", x))
  
  x2 <- x2 %>% # replace with original data frame
    left_join(parse, by = c("x2" = "full_name")) %>%
    rename(!!paste0(x.name, "_full_name") := "x2") %>% # operation for different input/output
    select(-llc, -suffix)

  return(distinct(x2)) # return, but remove dupliacte rows
}

# Example usage - separate names on an existing data set and join

#data2 <- data %>%
#  left_join(sep.names(data$name, "name_title"), 
#            by = c("name" = "x"))

# - Join with function on previous data frame
# - Name title will be the column name for the separate names (i.e. name_title_last_name)
# - Join on x and the original name column
```

each row should be a unique party at unique addresses in a unique eviction filing, but I noticed some repeat records that didn't seem to make any sense. I decided to extract only the unique recrods
```{r}
ev<-unique(courts_raw)
```

This cleans up the plaintiff and defendant addresses
```{r}
ev<-ev%>%
  mutate(party_address_before_space = gsub( " .*$", "", party_address_line1 ))%>% #first chunk of numbers or letters in party address before space to be used in the future to match with bldg_num 
  mutate(number_new = substr(party_address_line1,start=0,stop=1))%>% #first letter/number in the party address chunk 
  mutate(party_line1_num = ifelse((grepl("[[:digit:]]", number_new))==TRUE,"yes","no"))

ev<-ev%>%
  mutate(clean_party_address = ifelse(party_line1_num=="yes" & is.na(party_address_line3)==TRUE , paste(party_address_line1, party_address_line2), paste(party_address_line2,party_address_line3)))%>%
  mutate(party_first_three=substr(party_address_line1,start=0,stop=3))%>% #this will be used for PO boxes 
  mutate(last_2_party_char = str_sub(clean_party_address,-2)) #if the last two characters in the addressress are NA then it was an address that started with letters instead of numbers 
  
ev$clean_party_address<-ifelse(ev$party_address_before_space=="PO" | ev$party_first_three=="P O" , paste(ev$party_address_line1,ev$party_address_line2), ev$clean_party_address)
ev$clean_party_address<-ifelse(ev$last_2_party_char=="NA", paste(ev$party_address_line1,ev$party_address_line2),ev$clean_party_address)
ev$party_address_before_space<-gsub( " .*$", "", ev$clean_party_address )

ev<-ev%>%
  mutate(without_na=substr(ev$clean_party_address,1,nchar(ev$clean_party_address)-3))

ev$clean_party_address<-ifelse(ev$last_2_party_char=="NA", ev$without_na, ev$clean_party_address )


###once we're done with the clean party address fields we'll get rid of most of them

ev$number_new<-NULL
ev$party_line1_num<-NULL
ev$party_first_three<-NULL
ev$party_address_before_space<-NULL
ev$last_2_party_char<-NULL
ev$without_na<-NULL
```

```{r data wrangling}
# Wrangle into tidy data - every row should be a single observation (court case)

ev2 <- ev %>%
  mutate(case_filed_date = as.Date(case_filed_date, format = "%m/%d/%Y")) %>%
  mutate(judgment_date = as.Date(judgment_date, format = "%m/%d/%Y %H:%M")) %>%
  select(-(judgment_award_entered_date:satisfaction_type_desc), -(type_code:type_desc),
         -(base_type_code:base_type_desc), -district_desc, -county) %>% 
    # for now, remove judgements
    # also remove type and base_type (only 1) and county and district descriptions (may match on codes later)
  mutate(party_zip = as.integer(gsub(party_zip, pattern = "-[0-9]{0,4}", replacement = ""))) %>%
    # fix zip codes - first, remove trailing numbers on zip code
  mutate(party_zip = ifelse(nchar(party_zip) == 9, as.integer(substr(party_zip, 1, nchar(party_zip)-4)),
                            party_zip)) %>%
    # second, remove additional digits in zip if no dash is used
  mutate_all(funs(str_trim(.))) %>% # remove head and trailing spaces
  mutate(party_full_name2 = sapply(party_full_name, simpleCap)) %>% 
    # capitalize first letter of names, everything else to lowercaps
  select(current_case_number:party_full_name, party_full_name2, party_address_line1:judgment_type_desc) %>%
    # rearrange variables to keep names together
  group_by(current_case_number, case_party_involvement_code) %>% # group by case and party
  mutate(party_no = str_pad(as.character(row_number()), 2, pad = "0")) %>% 
    # number each party (i.e. first defendant, second defendant)
    # nested within str_pad which adds a leading zero (so 10 does not follow 1, etc.)
  mutate(def_no = ifelse(case_party_involvement_code == "DFD", n(), NA)) %>% # find number of defs and plas
  mutate(pla_no = ifelse(case_party_involvement_code == "PLN", n(), NA)) %>%
  unite(case_party_no, case_party_involvement_code, party_no) %>% # create variable for party and number
  group_by(current_case_number) %>%
  mutate(mult_judg_dates = n_distinct(judgment_date)) %>%
  mutate(mult_judg_dates = (ifelse(mult_judg_dates > 1, 1, 0))) %>%
  arrange(def_no) %>% mutate(def_no = def_no[1]) %>% # copy number of defs and plas to all rows
  arrange(pla_no) %>% mutate(pla_no = pla_no[1]) %>%
  ungroup() 

write_csv(ev2, "./post spreading evictions.csv") 

ev2<-ev2%>%mutate_all(as.character) %>%
  gather(variable, value, party_full_name:party_zip) %>%
  unite(temp, case_party_no, variable) %>%
  select(-case_party_involvement_desc) %>%
  spread(temp, value)
    # gather, unite, and spread: way to spread multiple columns, and make columns for the names and
    # addresses of each defendant and plaintiff


# clean variable types

ev3 <- ev2 %>%
  mutate_at(vars(case_filed_date), as.Date, format = "%m/%d/%Y") %>%
  mutate_at(vars(judgment_date), as.Date, format = "%Y-%m-%d") %>%
    # change both date variables to date
  mutate_at(vars(district_code:case_status_desc, judgment_type_code:judgment_type_desc), as.factor) %>%
    # change the others to factor
  mutate_at(vars(def_no:pla_no), as.numeric) %>%
    # change the number of defendants/paintiffs to numeric
  mutate(PLN_01_party_full_name = ifelse(is.na(PLN_01_party_full_name), PLN_02_party_full_name,
                                         PLN_01_party_full_name)) %>%
  mutate(PLN_01_party_full_name2 = ifelse(is.na(PLN_01_party_full_name2), PLN_02_party_full_name2,
                                         PLN_01_party_full_name2))
    # For over 200 cases with absent primary plaintiff names, use the second name in these cases

# separate data frame for judgments - also turn into wide data
ev2.judg <- ev %>%
  select(current_case_number, judgment_date:satisfaction_type_desc) %>%
  filter(!is.na(judgment_award_entered_date)) %>% # only consider rows with judgement awards (as date)
  group_by(current_case_number, award_party_role) %>%
  mutate(party_no = str_pad(as.character(row_number()), 2, pad = "0")) %>%
  mutate(cred_no = ifelse(award_party_role == "creditor", n(), NA)) %>%
  mutate(debt_no = ifelse(award_party_role == "debtor", n(), NA)) %>%
  unite(award_party_no, award_party_role, party_no) %>% # there are na party roles, have to remove later
  group_by(current_case_number) %>%
  arrange(cred_no) %>% mutate(cred_no = cred_no[1]) %>%
  arrange(debt_no) %>% mutate(debt_no = debt_no[1]) %>%
  ungroup() %>%
  mutate_all(as.character) %>%
  gather(variable, value, award_party_name) %>%
  unite(temp, award_party_no, variable) %>%
  spread(temp, value) #%>%
  #select(-(na_01_award_party_name:na_12_award_party_name)) # remove na columns

# Clean up variables
ev3.judg <- ev2.judg %>%
  mutate_at(vars(judgment_award_entered_date:judgment_award_satisfy_date), 
            as.Date, format = "%y-%m-%d %h:%m:%s") %>%
  mutate_at(vars(judgment_date), as.Date, format = "%y-%m-%d") %>%
    # change both date variables to date
  mutate_at(vars(judgment_type_code:judgment_type_desc,
                 judgment_award_property_type_code:judgment_award_property_type_desc,
                 satisfaction_type_code:satisfaction_type_desc), as.factor) %>%
    # change the others to factor
  mutate_at(vars(original_principal_amt, cred_no:debt_no), as.numeric)

# Note - we haven't done anything with this judgment data - so far it's removed so that it doesn't untidy the main eviction court data

write_csv(ev3, "./spread_clean_evictions.csv")
```


```{r}
### Standardize and match primary plaintiffs

# There are many different spellings of plaintiff names; there are also LLCs that are owned by the same property owner groups. This section matches based on misspellings, alternate spellings, and sources an external file where some cleaning was recorded.

# Since we are matching on address, make a single address column for DFD 1 and PLA 1
ev3.llord <- ev3 %>% # sample_n is more direct for rows than base R sample()
  mutate_at(vars(DFD_01_party_address_line1:DFD_01_party_zip,
                 PLN_01_party_address_line1:PLN_01_party_zip), funs(str_trim(.))) %>% 
  # remove trailing and leading whitespace
  mutate(dfd.zip = gsub(DFD_01_party_zip, 
                                   pattern = "-[0-9]{0,4}", replacement = "")) %>%
  # remove trailing numbers on zip code
  # mutate(dfd.zip = ifelse(nchar(as.integer(dfd.zip)) == 9, 
  #                        substr(zip, 1, nchar(dfd.zip)-4)), dfd.zip) %>%
  # also remove trailing numbers without dash
  mutate(dfd.address = paste(DFD_01_party_address_line1, DFD_01_party_address_line2,
                         DFD_01_party_address_line3, DFD_01_party_address_line4,
                         sep = ", ")) %>% # paste into single variable, separated by commas
  mutate(dfd.address = gsub("NA, |, NA", "", dfd.address)) %>% # remove NA (easier than filtering before)
  mutate(dfd.address = gsub("\\s\\,", "", dfd.address)) %>%
  # Do the same for landlord addresses
  mutate(pln.zip = gsub(as.character(PLN_01_party_zip), 
                                   pattern = "-[0-9]{0,4}", replacement = "")) %>%
  # mutate(pln.zip = ifelse(nchar(pln.zip) == 9, 
  #                        as.integer(substr(zip, 1, nchar(pln.zip)-4)), pln.zip)) %>%
  mutate(pln.address = paste(PLN_01_party_address_line1, PLN_01_party_address_line2,
                             PLN_01_party_address_line3, PLN_01_party_address_line4,
                             sep = ", ")) %>% 
  mutate(pln.address = gsub("NA, |, NA", "", pln.address)) %>% 
  mutate(pln.address = gsub("\\s\\,", "", pln.address)) %>%
  mutate(pln.address = ifelse(pln.address == "NA", NA, pln.address))
    # fix issue of NA characters in address

llc2 <- c("LLC|LLP|\\bINC\\b|\\bLP\\b|\\bL\\sP\\b|COMPANY|\\bCO\\b|\\bLTD\\b|
          \\bGP\\b|LIMITED PARTNERSHIP|PARTNERSHIP|LIMITED^|CORPORATION|\\bCORP\\b|INVESTMENTS|
          CONSULTING|OFFICE|APARTMENTS?|MORTGAGE|PROPERTIES|TRUST|NATIONAL ASSOCIATION|
          TOWNHOMES|ASSOCIATION|CONDOMINIUM|CONDO|OWNERS?|THE|CORP|
          GROUP|\\sL$|\\sL\\s|\\bN\\sA\\b") # Types of corporations or LLCs
# More inclusive - includes common words
library(humaniformat)
# separate names
ev3.llord1 <- ev3.llord %>%
  left_join(sep.names(ev3.llord$PLN_01_party_full_name, "llord"), # separate names
            by = c("PLN_01_party_full_name" = "x")) %>%
  select(current_case_number, dfd.zip:llord_suffix, PLN_01_party_city) %>% # select variables
  mutate_all(funs(toupper)) %>% # change to uppercase (all)
  mutate_all(funs(str_replace_all(., "[[:punct:] ]+", " "))) # remove punctuation
  #distinct(llord_full_name, .keep_all = TRUE) %>% # remove duplicates based on exact full name matching
    # don't want to remove duplicates? - this hinders matching same or similar names

# remove common LLC and business titles/termms
ev3.llord2 <- ev3.llord1 %>%
  mutate(llord_full_name2 = str_replace_all(llord_full_name, llc2, "")) %>% 
    # remove LLCs, Corp, Bank, etc. (i.e. don't match on this)
  mutate(llord_full_name2 = str_trim(llord_full_name2, side = "both")) %>% 
    # remove extra spaces
  rename(prop_address = dfd.address,
         llord_address = pln.address)
    # rename variables

# See most common names & corrections
#ev3.llord.test2 <- ev3.llord2 %>%
#  mutate(match = str_detect(llord_full_name, llc2)) %>%
#  mutate(llord_full_name2 = str_replace_all(llord_full_name, llc2, "")) %>%
#  group_by(llord_full_name, llord_full_name2, match) %>%
#  summarize(n = n()) %>%
#  arrange(desc(n))

# Just match based on name
ev3.llord3 <- ev3.llord2 %>%
  mutate(block = ifelse(!is.na(llord_first_name) & !is.na(llord_last_name), 
                        paste0(substr(llord_last_name, 1, 3), 
                              str_extract(llord_address, "\\d+"), sep = ""),
                        # TODO : only capture first match (numbers, whether address or PO Box)
                        paste(substr(llord_full_name2, 1, 3),
                              str_extract(llord_address, "\\d+"), sep = ""))) %>%
    # block: first digits of last name or full name (if entity)
    # also first block of numbers (so address number or PO Box number, for example)
  mutate(block = str_replace_all(block, "\\D[NA]$", "")) %>% # remove NAs
  mutate(block = str_replace_all(block, " ", "")) # remove spaces

ev3_backup<-ev3

#get rid of other address fields besides the property fields and the landlord address fields 
ev3 <- ev3%>%
  select(-(14:427))

ev3 <- ev3 %>%
  left_join(ev3.llord3, by=c("current_case_number"="current_case_number"))

```

#this creates a simplified version of the judgement decision categories 
```{r}
simple_judgments<-read_csv("./simple_judgements.csv")

ev3<-ev3%>%
  left_join(simple_judgments, by=c("judgment_type_desc"="judgment_type_desc"))
```

creating a useable judgement date field that can be plotted and filtered correctly
```{r}
ev3<-ev3%>%
  mutate(judgement_mdy = gsub( " .*$", "", judgment_date ))

ev3$judgement_ymd<-ymd(ev3$judgement_mdy)

ev3<-ev3%>%
  mutate(judgement_year = year(judgement_ymd))#just a column for year
```

Joinging the county names the counties back to the new wide evictions data frame
```{r}
ev_county<-ev%>%
  select(current_case_number,county)%>%
  group_by(current_case_number,county)%>%
  summarise(total=n())

ev_county$total<-NA

ev3<-ev3%>%
  left_join(ev_county, by=c("current_case_number"="current_case_number"))

```


Unique cases per year total and by county, including judgements
```{r}
cases_yearly<-ev3%>%
  select(current_case_number, judgement_year, county, judgement_type_coll,judgement_ymd)

cases_yearly<-unique(cases_yearly)
```

```{r}
mpls_zips<-read_csv("./mpls_zips.csv")

mpls_zips$dfd.zip<-as.character(mpls_zips$dfd.zip)

mpls_evs<-ev3%>%
  left_join(mpls_zips, by=c("dfd.zip"="dfd.zip"))%>%
  filter(judgement_type_coll=="Eviction",minneapolis=="minneapolis")%>%
  filter(judgement_year==2018|judgement_year==2019|judgement_year==2017)%>%
  group_by(prop_address,dfd.zip,minneapolis)%>%
  summarise(total_evs=n())


write_csv(ev3,"evictions_cleanLLname_join.csv")
write_csv(mpls_evs, "evictions_in_mpls_2017_2019.csv")
  #put in a grepl about the clean address containing minneapolis somewhere in there

  #append the dfd blank add to the henn county one after its been grouped by address and total number of evictions
  
  #more than one eviction? write it out, geocode it, and then spatially join it to the rental license data. Maybe just do in the past year for evictions.....


```

